---
title: "05-heatmap-both-datasets"
author: "Ariel Marcy"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, results='hide')
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Heatmaps both with-allometry and without-allometry 3D coordinates
1) Make heatmaps with shape dataset including allometry.
        A) Smallest to mean to largest
        B) Smallest projected shape to mean to largest projected shape
        C) PC1 minimum to mean to PC maximum

Produces multipanel Figure 2.

2) Make heatmaps for residuals aka size-less shape dataset.
        A) PC1 min to mean to PC1 max
        B) PC1 min on subset without Notomys to mean to PC1 max of this subset
        C) PC2 min to mean to PC2 max

Produces multipanel Figure 3.

### Load packages and data
Including installing libraries for landvR package, which creates the heatmaps.
```{r message = FALSE}
library(geomorph) # CRAN v4.0.6
library(devtools) # CRAN v2.4.5
if(!require(dispRity)) install.packages("dispRity")  
if(!require(landvR)) install_github("TGuillerme/landvR")  
library(landvR)   # [github::TGuillerme/landvR] v0.5.2
library(png)      # CRAN v0.1-8 
library(Rvcg)
# This sets the wd to local
library(rstudioapi) # Safely Access the RStudio API
setwd(dirname(getActiveDocumentContext()$path))

load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/04-PCA_data.rda")
load(file = "../Data/Processed/allometry_residuals.rda")
```

## 1) Create heatmap figures
This script is dominated by the code needed to generate, export, and clean the heatmap images which are tiled in Figures 2 and 3. 

### Make heatmaps involving the smallest and largest native species 
We specify native species because technically *Mus musculus* has smallest centroid size but since it is not native, it is less relevant than *P. delicatulus*, which is only slightly larger than it.
```{r}

# Get index for smallest and largest native species 
sm.index <- which(info.means$Species == "del")
big.index <- which(info.means$MeanCsize == max(info.means$MeanCsize))

# Designate mean shapes for individual species with indices
smallest.mean <- mean.shapes[, , sm.index]
biggest.mean <- mean.shapes[, , big.index]
#which species is that?
dimnames(mean.shapes)[[3]][ big.index]


# Measure spherical coordinates differences from the smallest to the biggest mean. This results in a list of just 1 (diff.mean[[1]])
diff.mean_smallest_largest <- coordinates.difference(coordinates = biggest.mean, reference = smallest.mean, type = "spherical")


```




### Make heatmaps for PC1 with residual coordinates
Steps as above.
```{r}
# Find mean configuration - i.e. mean shape in dataset
res.mean <- (select.procrustes(residuals_lm, selector = base::mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean <- coordinates.difference(coordinates = residuals_lm, reference = res.mean, type = "spherical")

# Ordinate the data
twoD.shape.r <- two.d.array(residuals_lm)
ordination.r <- stats::prcomp(twoD.shape.r)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- residuals_lm
gpagen.shape$consensus <- res.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1. VW I added "coords" here because, as per explanation in the help file, this should give us the differences between the specimen with the minimum and maximun variation and gpagen.shape has $consensus which we don't want.
PC1.var.r <- variation.range(gpagen.shape$coords, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r <- gm.prcomp(residuals_lm)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h <- wrap.PCA.r$shapes[[1]]$min
res.PC1.max.h <- wrap.PCA.r$shapes[[1]]$max
```

### Make heatmaps for PC2 residual coordinates
Steps as above but for PC2; hypothetical shapes
```{r}
# Measure extremes of variation from mean on PC1
PC1.var.r2 <- variation.range(gpagen.shape, return.ID = FALSE, axis = 2, ordination = ordination.r, type = "spherical")

# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.h <- wrap.PCA.r$shapes[[2]]$max  # switch for consistency
res.PC2.max.h <- wrap.PCA.r$shapes[[2]]$min

# Calculate distances between shapes to ensure correct point color
min.coords <- cbind(res.mean, res.PC2.min.h)
min.dist <- apply(min.coords, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
max.coords <- cbind(res.mean, res.PC2.max.h)
max.dist <- apply(max.coords, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
```

### Export the variation for shape residual change
```{r}
open3d( windowRect = c(0, 0, 1200,700) )
mfrow3d(nr=1,nc=2)

#open3d(windowRect = c(0, 0, 2400,2400))
procrustes.var.plot(res.PC1.max.h,res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90, zoom=0.7)

next3d()
procrustes.var.plot(res.PC2.max.h,res.PC2.min.h, col = heat.colors, col.val = min.dist)
view3d(phi = 90, zoom=0.7)
rgl.snapshot("../Data/Processed/Heatmaps/LM_res_heatmaps.png")


# Remove white space from all heatmaps (requires ImageMagik installed on PC; when installing ImageMagick, make sure you tick the option of installing legacy utilities, otherwise the below won't work)
system("mogrify ../Data/Processed/Heatmaps/*.png -trim ../Data/Processed/Heatmaps/*.png")

```
